Governing through patronage: the bargain for education in decentralized Brazil

### 2020-09-10 paper/paper.tex
include table

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,      
    urlcolor=blue,
    citecolor=blue
}

@@

\begin{figure}[h]
  \includegraphics{plots/plot_coalition_coef_mun.pdf}
  \caption{{\bf Executive coalitions and staff turnover.} An increase in the share of legislative seats held by the mayoral coalition decrease the amount of turnover for teachers and school principals, including hires or dismissals. Model 1 presents results on the turnover index. Models 2 and 3 are logistic regressions at the individual, bureaucrat level. Models 4 and 5 are aggregated at the municipal level. where the outcome of interest is the proportion of staff either hired or dismissed at a given year. Models 1, 3, and 5 include year and state fixed effects.}
  \label{fig:coalition_turnover}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics{plots/plot_interactions_logit}
    \caption{Predicted values for bureaucratic turnover. The graph plots the proportion of educational staff either hired or fired for any given year against the share of seats controlled by the executive. The circles' size is proportional to the number of observations contained at each value of the legislative share of seats.}
    \label{fig:turnover}
\end{figure}

Finally, I demonstrate that

Moving on to characterizing the downstream effects of patronage, I find that teacher and school principal turnover have significant, negative effects on average test scores. These results are robust to alternative specifications of turnover, as well as the use of different quality metrics (SAEB, SPAECE) in our estimations. Table \ref{fig:hlm_mods} presents the results of the estimation on SAEB test scores. We present two specifications for turnover. Turnover index measures the amount of turnover in teachers at the school level. Work experience for teachers and school principals serve as an alternative measure of turnover. Personnel who has only recently entered the school recently amass a shorter amount of work experience (less than two years).

@@

% As a final note, this set of estimations are based on observational data and therefore suffer from well-known concerns of endogeneity. It is in these circumstances that validating causal mechanisms through fieldwork and careful data analysis can increase the validity of a causal claim. In-depth interviews with local actors involved in managing education, as well as the use of different measurements for both educational staff turnover and student learning, provides strong evidence that staff turnover has a negative impact on student learning. Bureaucratic turnover stems from political considerations.

>>>>

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,      
    urlcolor=blue,
    citecolor=blue
}
\captionsetup[figure]{font=small,labelfont=small}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\linewidth]{plots/plot_coalition_coef_mun.pdf}
    \caption{{\bf Political alignment and staff turnover.} An increase in the share of legislative seats held by the mayoral coalition decrease the amount of turnover for teachers and school principals, including hires or dismissals. Model 1 presents results on the turnover index. Models 2 and 3 are logistic regressions at the individual, bureaucrat level. Models 4 and 5 are aggregated at the municipal level. where the outcome of interest is the proportion of staff either hired or dismissed at a given year. Models 1, 3, and 5 include year and state fixed effects.}
    \label{fig:coalition_turnover}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\linewidth]{plots/plot_interactions_logit}
    \caption{Predicted values for bureaucratic turnover. The graph plots the proportion of educational staff either hired or fired for any given year against the share of seats controlled by the executive. The circles' size is proportional to the number of observations contained at each value of the legislative share of seats.}
    \label{fig:turnover}
\end{figure}

Finally, I show that the executive-legislative bargain extends to second-term mayors. I subset the data used for the above analysis to only mayors who are reelected, and re-run the previous analyses with the same specification. I find that the results are similar in magnitude and precision to those presented above. This finding suggests that patronage is not motivated solely by reelection concerns. Rather, patronage achieves an important policy goal for incumbent mayors who seek to implement their preferred policy, regardless of whether they are just initiating their mandate or concluding it.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\linewidth]{plots/plot_coalition_coef_mun_second.pdf}
    \caption{\bf{hello}}
\end{figure}

Moving on to characterizing the downstream effects of patronage, I find that teacher and school principal turnover have significant, negative effects on average test scores. These results are robust to alternative specifications of turnover, as well as the use of different quality metrics (SAEB, SPAECE) in our estimations. Table \ref{tbl:student_learning} presents the results of the estimation on SAEB and SPAECE test scores. We present two specifications for turnover. Turnover index measures the amount of turnover in teachers at the school level. Work experience for teachers and school principals serve as an alternative measure of turnover. Personnel who has only recently entered the school recently amass a shorter amount of work experience (less than two years).

\begin{landscape}
    \begin{table}[t]
      \centering
      \footnotesize
      \input{results/student_learning.tex}
      \caption{{\bf Bureaucratic turnover and student learning} Teacher and school principal turnover have a negative effect on student learning. Models 1 and 2 present results for teacher turnover index constructed at the school level. Models 3 and 4 estimate the effect of new teachers and school principals entering the school (less than two years). All models include year and state fixed effects.}
      \label{tbl:student_learning}
    \end{table}
\end{landscape}

@@

As a final note, this set of estimations are based on observational data and therefore suffer from well-known concerns of endogeneity. It is in these circumstances that validating causal mechanisms through fieldwork and careful data analysis can increase the validity of a causal claim. In-depth interviews with local actors involved in managing education, as well as the use of different measurements for both educational staff turnover and student learning, provides strong evidence that staff turnover has a negative impact on student learning. Bureaucratic turnover stems from political considerations.
